\documentclass[oldfontcommands]{ucbthesis}
\usepackage[backend=biber]{biblatex}
\usepackage{url}

\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{blindtext}
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={Fabbed to Sense : a thesis by Valkyrie A. Savage},
pdfauthor={Valkyrie A. Savage},
pdfkeywords={},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% Double spacing, if you want it.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\bibliography{references}

\hyphenation{mar-gin-al-ia}

\begin{document}
\input{inputmacros}

% Declarations for Front Matter

\title{Fabbed to Sense: Co-Design of Geometry and Sensing Algorithms for Interactive Objects}
\author{Valkyrie Savage}
\degreesemester{Fall}
\degreeyear{2015}
\degree{Doctor of Philosophy}
\chair{Professor Bj\"orn Hartmann}
\othermembers{Professor Carlo S\'equin \\
Professor Eric Paulos\\
Professor Paul Wright}
\numberofmembers{4}
\prevdegrees{B.A., B.S. (Indiana University Bloomington) 2010}
\field{Computer Science}
\campus{Berkeley}

\maketitle
\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
\begin{center}
Dedicated to the adventure-hungry:\\\vspace{12pt}
whimsy, discovery, and enchantment are out there.
\end{center}
\vfil\null
\end{dedication}

\tableofcontents
\clearpage
\listoffigures
%\clearpage
%\listoftables

\begin{acknowledgements}
Many, many thanks are due the committee for their dedication and time in helping guide me through both the preparation of this document and the odyssey of graduate school itself. Special thanks are due, naturally, to Bj\"orn, for taking a chance on a grad student who clearly had no idea what she was doing. He has been endlessly patient in allowing me to explore new things and experiment with crazy ideas, as well as tolerating my occasional jaunts into absurd extracurriculars. It has been a wild ride, and you are a wonderful mentor! 

My collaborators at various times, including Berkeley grad students Andrew Head, Daniel Lim, and Michelle Nguyen, and undergraduates Xiaohan Zhang, Colin Chang, Jingyi Li, Eldon Schoop, and Mitchell Karchemsky; Autodesk researchers Ryan Schmidt, Tovi Grossman, and George Fitzmaurice; Adobe researchers Dan Goldman, Gautham Mysore, and Wilmot Li; and MIT-student-cum-Stanford-professor Sean Follmer; have aided me at every turn. Without them this work would naturally not have been possible. There are even more of you that I never ``officially'' collaborated with, in the BiD lab and outside it, that gave me feedback and occasionally beer. Thanks to Shiry Ginosar, Celeste Roschuni, Mark Fuge, Lora Oehlberg, Wes Willett, Andy Carle, Drew Sabelhaus, Jer Faludi, Armin Samii, Florian Altvater, Mark Oehlberg, Chris Myers, Drew Fisher, Peggy Chi, Laura Devendorf, Cesar Torres, Tim Campbell, and Joanne Lo for being awesome. I appreciate all your advice on ideas, implementation, and life.

I also owe a karmic debt to all Juggers, from the Berkeley Riot and elsewhere, with whom I have interacted over the past two years. Lacking a community with such a sense of play, I doubt I would have maintained my sanity long enough to complete this opus.

Finally, I thank my parents (Howard and Angela Savage), my sister (Venus Savage), and my husband (Evan Savage) for believing in me. I daren't try elaborating on all the amazing things you do, but thank you so much. $\heartsuit$

Now, because this is my thesis and I can do what I want, here's a cowsay\footnote{\url{https://en.wikipedia.org/wiki/Cowsay}}:
\begin{verbatim}
     _________________________________________
    / thank you to everyone; this has been an \
    \ amazing journey!                        /
     -----------------------------------------
            \   ^__^
             \  (oo)\_______
                (__)\       )\/\
                    ||----w |
                    ||     ||
\end{verbatim}
\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\include{01introduction}
\include{02fabrication}
\include{03relatedwork}
\include{04midas}
\include{05lamello}
\include{06sauron}
\include{07adiscussion}
\include{07bfuturework}
%\include{08conclusion}

% \appendix

\printbibliography

\end{document}