% !TEX root = thesis.tex
\chapter{Introduction}

\begin{quote}
Design is not just what it looks like and feels like. Design is how it works.

--- Steve Jobs
\end{quote}

Our environment is replete with products that have dedicated physical user interfaces like game controllers, musical instruments or personal medical devices. While the ubiquity of smart phones has led to a rise in touchscreen applications, retaining physicality has important benefits such as tactile feedback and high performance manipulation \cite{klemmer-bodies}. For example, gamers prefer physical input for speed and performance, musicians for virtuosity and control. Rapid additive manufacturing technologies enable designers and makers (henceforth we refer to both groups jointly as ``makers'') to quickly turn CAD models of such future devices into tangible prototypes. While such printed form prototypes can convey the look and feel of a physical device, they are fundamentally passive in that they do not sense or respond to manipulation by a user. Building integrated prototypes that also exhibit interactive behavior requires adding electronic sensing components and circuitry to the mechanical design.

Existing research has developed electronic toolkits that lower the threshold of making physical prototypes interactive \cite{arduino,greenberg-phidgets}. However, such toolkits still require makers to manually assemble printed parts and sensors. Such assembly may also require significant changes to a 3D model (e.g., to add fasteners or split an enclosure into two half shells). Detailed electro-mechanical co-design is time-consuming and cumbersome and mismatched with the spirit of rapid prototyping. Alternatively, makers may instrument the environment with motion capture \cite{akaoka-displayobjects} or depth cameras \cite{wilson-depth} to add interactivity, but these approaches limit makers to testing prototypes inside the lab in small, restricted areas.

\emph{We aim to uncover the most cost-effective, fast, and flexible ways of sensing rapid-prototyped input devices.} This suggests several requirements:
\begin{enumerate}
\item \textbf{cost-effective} : neither the fabrication means and techniques required for sensing nor the sensor itself should be prohibitively expensive to makers, and environmental sensors (e.g., those in an average smart phone) should be leveraged where possible. We focus on \emph{single-sensor} techniques, where sensing of multiple user inputs can be achieved by affixing a sensing apparatus to a single point. \valkyrie{those aren't exactly the same; reword}
\item \textbf{fast} : fabrication and assembly of senseable devices should not take significantly longer than comparable passive devices, and necessary digital modifications should be performed automatically when complex or be reduced to templates when simple.
\item \textbf{flexible} : the means of sensing a prototype object should not impose undue burden on the possible designs of that object. Sensing techniques should accommodate a wide variety of input types (e.g., buttons, sliders, and dials) and body types (e.g., convex, concave, 3D).
\end{enumerate}
We propose a novel way of ensuring these properties: users create and modify digital design files, then fabricate them using digital fabrication machines. This creates a \emph{link between the digital and physical models}, which we leverage to inform and improve our sensing capabilities.

\section{Contributions}

This thesis explores the realm of physically prototyping tangible input devices using digital fabrication machines. We have built several prototype design and sensing systems designed to test different parts of the design space. Thus, this thesis makes the following contributions:

\begin{enumerate}
\item \valkyrie{a summary of today's fabrication machine capabilities and how they tie to sensing systems?}
\item Midas, a method to create custom-shaped, flexible capacitive touch sensors by synthesizing sensor pads and auto-routing connections, as well as instructions for assembly and use, from a high-level graphical specification \begin{enumerate}
    \item a design tool using this method to enable users to to fabricate, program, and share touch-sensitive prototypes
    \item an evaluation demonstrating Midas’s expressivity and utility to designers
    \end{enumerate}
\item Lamello, a technique which integrates algorithmically-generated audio-producing tine structures into movable components, creating passive tangible inputs that do not require training examples for accurate sensing. \begin{enumerate}
    \item a design pipeline which predicts tine frequencies (and an evaluation that they can be accurately predicted) and senses user manipulation of components in real time
    \item a discussion of information encoding techniques useful for this technique \valkyrie{not really a contribution? just a summary... should this contribution be a library of parts, instead?}
    \end{enumerate}
\item Sauron, a design tool enabling users to
rapidly turn 3D models of input devices into interactive 3D printed prototypes where a single camera senses input \begin{enumerate}
    \item a method for tracking human input on physical components using a single camera placed inside a hollow object
    \item two algorithms for analyzing and modifying a 3D model’s internal geometry to increase the range of manipulations that can be detected by a single camera.
    \item an informal evaluation of our implementation of these techniques usable on models constructed in a professional CAD tool.
    \end{enumerate}
\end{enumerate}

\section{Dissertation Sketch}

This section presents a brief sketch of the structure of this dissertation by chapters.

This dissertation first investigates the capabilities of modern digital fabrication machines and discusses single-sensor techniques compatible with those capabilities. Second, a solid foundation of related work is laid out. We then discuss three instances of analysis and design tools that help users design, fabricate, assemble, and \emph{use} input devices sensed in a variety of ways; for each technique we demonstrate its flexibility for use in many types of input devices. Finally, we conclude with suggestions for future work.

\subsection{Fabrication \& Sensing (Chapter 2)}

This chapter explores modern digital fabrication machines and the properties that can be employed in objects they fabricate. We discriminate properties that are commonly available in hobbyist machines from those that require a high-end professional machine.

Chapter 2 further discusses single-sensor techniques compatible with those capabilities, as well as what makes each technique promising or challenging. This thesis ultimately selects a few points to further examine in this space.

\subsection{Related Work (Chapter 3)}

We situate this thesis in the realm of what's been done. In general, we draw heavily on work from four major traditions: graphics and optimization, sensing digitally-fabricated devices, modeling 3D objects, and creating prototypes.

\subsubsection{Graphics/Optimization}
The graphics community embraced digital fabrication early: they provide some of the earliest research on using pre-fabrication simulation to control post-print deformation behaviors \cite{bickel-deformation} and appearance \cite{lan-appearance}. We, too, perform pre-print simulation and optimization of 3D models, but for the purpose of creating \emph{interactive} objects.

\subsubsection{3D design}
Many professional tools for 3D modeling exist \cite{solidworks, rhino}, and they serve their target users well. Researchers have made significant strides in inventing new styles of more accessible interactions for 3D modeling, for example by capturing toolpaths \cite{willis-interactive} or clay models \cite{savage-mmarks}. For the purposes of our investigations in this thesis, we created design tools to help author objects compatible with our sensing techniques: they are a complement to, rather than a replacement for, existing design tools research.

\subsubsection{Sensing Digitally-Fabricated Devices}
Sensing forms a key component of interactive devices, and thus has been significantly explored in the past. One common technique for sensing objects uses machine learning and guided manual training to detect interactions using sound \cite{ono-touchandactivate,laput-acoustruments}, capacitance \cite{sato-touche} or other signals. Our techniques focus on sensing done without machine learning, and often without any training at all. Inspired by ``Sensing Through Structure'' \cite{slyper-structure}, we design objects whose properties we know, which can inform how we sense them.

\subsubsection{Prototyping}
Abundant research has examined questions around the types of prototypes that designers build in the course of designing a new object \cite{houde-prototypes}. Other work looks further into more facile ways of creating electronic and functional, for example using snap-together circuits \cite{littleBits, hartmann-dtools, villar-gadgeteer} or smart circuit substrates \cite{villar-voodooio}. One important limitation of these investigations is that they are limited to sensors that \emph{already exist}: designers must make do with what they can buy. We conversely focus on customizable inputs that can be configured exactly as a designer wants them. Once an object is assembled, its functionality must be defined. We leverage techniques like programming by demonstration \cite{myers-pbd, hartmann-dtools} to help users define their objects' interactivity.

%% \valkyrie{bjoern's these are really long... how long do they need to be??}

\subsection{Midas: Capacitive Sensing of Custom 2D Layouts (Chapter 4)}

Our first exploration examines fabrication of 2D conductive materials sensed capacitively. Midas pushes towards a world in which prototyping physical interactive devices is as easy as prototyping GUI devices. Midas's sensing is capacitive, building on the popular capacitive touch sensing built in to smart phones and other touch-screen devices, but it explores how to prototype touch-based interactions where input and output are \emph{not} co-located, as they are on touch screens. Designers are given a drag-and-drop authoring system to create capacitive touchpads on the surface of objects, and from these designs generates 2D design files for fabrication. These files can then be cut from a conductive material and sensed using an automatically-configured microcontroller board. Midas also offers support for programming the input devices designed via PBD and websockets.

The advantages of capacitive sensing in this manner are numerous. It can be deployed on any flat, ruled, or developable object's surface. The sensors are cheap and easy to fabricate---whether on a vinyl cutter, using a circuitboard mill, or in inkjet printed conductive ink. Sensor assembly is \emph{fast}: it is color-coded for assembly, and users simply need to attach (e.g., using tape) the wires to their fabricated sensor pads.

In this chapter, we will detail Midas's implementation, as well as discussing possible uses, drawbacks of the current system, and work for the future.

\subsection{Lamello: Acoustic Sensing of 2D/3D Mechanisms (Chapter 5)}

Second, this thesis dives into an examination of acoustic-based sensing for devices fabricated in 2D, 3D, or a combination. The Lamello project investigates the use of tine-like structures for repeatable and predictable audio frequency generation. These tines can be printed at interaction points (e.g., under the path of a human input slider) such that they are struck when a user manipulates input components. The mechanical vibrations created by striking the tines can be detected with a contact microphone and classified using frequency analysis.

Leveraging uniform 2D lasercut or 3D printed plastics as sound-creating input devices offers flexibility to designers with different levels of access to fabrication machines. In addition, beyond Midas's offering of flat input surfaces activated by a simple touch, Lamello explores input mechanisms that users can push, slide, and turn. The technique of passive audio generation for sensing also opens up opportunities in the future Internet of Things: multiple unpowered Lamello-type input devices may be placed in the environment and sensed by a single microphone, perhaps located on a laptop or smartphone.

This chapter details our experiments confirming that our 3D-printed tine structures behave in predictable ways in spite of the non-uniform nature of the materials that comprise them. We also discuss, using several exemplars, techniques for integrating the tines into existing input component designs. Further, we describe information encoding principles for tine generation.

\subsection{Sauron: Vision-Based Sensing of 3D Printed Mechanisms (Chapter 6)}

Finally, we explore full 3D input devices sensed using computer vision. Sauron is a design and sensing toolkit for creating 3D printed input devices---which can include components like joysticks or dials---sensed with a single embedded camera. The Sauron tool makes automatic modifications to allow for this sensing, reconfiguring the \emph{interior} parts of the inputs and performing interference simulation.

Sauron's interfaces have additional flexibility over those for Midas or Lamello: they allow \emph{continuous} sensing of user input. They can be fabricated on any 3D printer, and the pre-fabrication simulation process relies only on geometry rather than any particular materials properties for its processing. \valkyrie{not sure what else to put here...}

In this chapter we describe our implementation of Sauron as a plugin for a commercial CAD tool, as well as the vision sensing code we built. Finally, we elaborate on limitations of the current system and places we may improve it, as well as future work in the area.

\subsection{Conclusion \& Future Work (Chapter 7)}

The final piece of this thesis reviews the contributions described and re-evaluates assumptions made in the projects constituting its main chapters. These re-evaluations lead to interesting pointers for future work in cooperation among fabrication machines, branching prototypes, and machine-optimized prototype designs.

\section{Statement of Multiple Authorship and Prior Publication}

The research presented in this dissertation was not undertaken by me alone. While I initiated
and led all projects described herein, I must acknowledge the contributions of my talented group of collaborators: without their efforts, this research could not have been realized in its current scope.

In particular, Midas's routing features were implemented by Xiaohan Zhang, and the video was created by Lora Oehlberg. Andrew Head performed much debugging and audio testing on the Lamello project, and that project benefited from the wisdom of my collaborators Dan Goldman (who provided the initial idea), Gautham Mysore, and Wilmot Li at Adobe. Sauron's computer vision was implemented by Colin Chang, and many thanks are due Mark Oehlberg for assisting in the creation of the necessary circuitboards for that project.

My advisor, Bj\"orn Hartmann, provided invaluable advice and guidance on all projects detailed in this document.

This dissertation is partially based on papers previously published in ACM conference proceedings; I am the primary author on all publications. In particular, Midas was published at UIST 2012 \cite{savage-midas}; Lamello at CHI 2014 \cite{savage-lamello}; and Sauron at UIST 2013 \cite{savage-sauron}.